Discriminative Sensing Based on Signal Processing for Information Security Analysis

### 2017-11-10 docs/thesis/chapters/3_mobile/mobile.tex
mobile chapter

This work proposes an archtecture and approach for user behavior analysis based on Model Order Selection (MOS) \cite{tenorio2013greatest}, in order to detect possible threats, and to reduce the risks and the harm of the most common threats, which are the expired user misusing password and the intruder attack. The key expiration period is safely incorporated into the proposed system solution in order to enhance security. Additionally, the behavioral analysis can indicate well known malicious behaviors, their variations, as well as novel attacks, which present low or high variance in comparison to legitimate user behaviors. The main target of our proposed solution is to provide a maximum defense at the minimal resource cost.

The work is structured as follows. Section \ref{sec:3_related_works} analyzes the most common security problems in the mobile cloud environment and the solutions for offline protection in the BYOD world. Section \ref{sec:3_the_client_security_architecture} outlines the mobile security architecture of the proposed solution. Section \ref{sec:3_offline_mode} presents the detailed scheme of the proposed solution to the problem of offline mobile client security. Section \ref{sec:3_algorithms} explains the algorithms for MOS. Section \ref{sec:3_results} discusses the common threat scenarios, the data modeling, the complexity analysis and dicussion of the practical implementation results. Section \ref{sec:3_conclusion} concludes the chapter.

@@

In many cases the industrial providers of the mobile security API avoid the problem by completely forbidding the offline access to the protected app, i.e. SAP Mocona, which operates as a secure client wrapping layer \cite{Lelyveld2013}. Due to various constraints such as traffic load, travelling and ease of access, the SME business procedure may require such access. To the best of our knowledge, the offline mode security problem has not yet been deployed, neither in academia nor in the industry \cite{yovel2014, khan2015cloud, khan2014survey, khan2013towards}. Therefore, the main concern of this proposal is the protection of the mobile client and its data in offline mode, when the functions of data protection cannot be offloaded to a cloud or a trusted party.

@@

The core set of functions and protocols can be divided into three sets of operations as shown in Figure \ref{fig:3_01}.

@@

Figure \ref{fig:3_02} depicts the client-side protection mechanisms. The client should support 4 subsystems: i) encryption subsystem that contains the procedures of encryption and decryption; ii) storage subsystem that provides the downloaded shares and key storage protection; iii) threat intelligence unit that provides the constant monitoring, and iv) the communication subsystem. In short, all security procedures are connected to 4 groups of operations: file request and receiving; encryption and decryption; file and key storing; monitoring and analysis.

@@

This architecture consists of the modules of cryptographic functions, threat intelligence infrastructure, communication with server and storage. This work focus on the threat intelligence infrastructure module.

The threat intelligence infrastructure takes into account simple actors such as the time counter for the key expiry period, the counter of unsuccessful tries in order to protect from brute force attacks, and MOS-inspired statistics analyzer. Functions such as alerting and deleting the expired key belong to this block as well. These functions are described in Subsection \ref{sec:3_mos}.

The communication with server includes the separate sender and receiver to check the user credentials and receive new keys. The remaining data like access list and files can be sent via unprotected channels.

@@

In the online mode the mobile client uses the secure communication with the server in order to verify the validity of user’s credentials. On the contrary, the offline protection model should be approached independently. The proposed architecture for offline mobile security is represented in Figure \ref{fig:3_03}.

@@

The MOS schemes can rely on sample covariance of zero mean variables (called as zero mean covariance for the sake of simplicity) and sample covariance of zero mean and unitary standard deviation (called as zero mean and standardized covariance for the sake of simplicity) variables, where the former is useful to identify abnormalities caused by large amounts of operations during a period, while the latter is applied to identify anomalies on sparse or subtle number of file operations.

@@

Once the $\mathbf{\hat{R}}_{yy}$ or $\mathbf{\hat{R}}_{zz}$ have been obtained for MOS in order to anomaly detection, for the sake of simplicity, we refer to $\mathbf{\hat{R}}_{yy}$ or $\mathbf{\hat{R}}_{zz}$ as a matrix $\mathbf{C}$. Therefore, the next step of the algorithm is the eigenvalue decomposition (EVD), calculated according to $\mathbf{C}^{(q)} = \mathbf{V}^{(q)}\mathbf{\Lambda}^{(q)}\mathbf{V}^{(q)^{\rm T}}$, in order to obtain the vector of eigenvalues $e$, as following:

\begin{equation}\label{eq:3_eq07}
	\mathbf{e}^{(q)} = \rm diag(\mathbf{\Lambda}^{(q)}),
\end{equation}

@@

Note that some MOS schemes may also require the amount of time that compose a time frame, such as $\hat{d} = \textrm{MOS}(\boldsymbol{e}_{\rm max}, M)$. For more information about MOS schemes, interested readers are referred to \cite{da2009comparison}.

@@

This section provides the detailed analysis of the results fro the proposed approach regarding security and complexity analysis.

@@

The session time defines the period when operations can be performed until the next session renewing. During this period, it is still necessary to identify attacks and malicious behavior on file operations, in order to avoid fast attacks to perform unauthorized access to information or data modification. Some attacks present behavioral patterns based on abrupt number of operations, such as the ransomware attack, which is a growing attack \cite{McAfee2015} that blocks the access to valuable resources and requires a payment in order to unblock the content. The access to the resources can be blocked by the attacker through some techniques, when the content is encrypted by the attacker, the ransomware attack can be called cryptoransomware \cite{kaspersky2014}.

@@

\begin{enumerate}
	\item file decryption; 
	\item decrypted file reading;
	\item decrypted file execution. 
\end{enumerate}

@@

\begin{enumerate}
	\item file encryption;
	\item decrypted file writing.
\end{enumerate}

@@

Table \ref{tab:3_04} presents the data modeling time and the processing time of eigenvalues decomposition calculations to be applied to anomaly detection from user operation logs of Storgrid mobile client. The information presented by column are the device model, the log size in megabytes, the window size in minutes, the data modeling time in milliseconds, the average of eigenvalue decomposition time in milliseconds, the standard deviation of eigenvalue decomposition time in milliseconds, the minimum of eigenvalue decomposition time in milliseconds and the maximum of eigenvalue decomposition time in milliseconds.

@@

@@

Future works in the area can further explore enhancements in the analytics methods as well as to extended the approach to be used by mobile devices with even more severe resources constraints.

>>>>

This work proposes an archtecture and approach for user behavior analysis based on Model Order Selection (MOS) \cite{tenorio2013greatest}, in order to detect possible threats, and to reduce the risks and the harm of the most common threats, which are the expired user misusing password and the intruder attack. The key expiration period is safely incorporated into the proposed system solution in order to enhance security. Additionally, the behavioral analysis can indicate well known malicious behaviors, their variations, as well as novel attacks, which present low or high variance in comparison to legitimate user behaviors.

The work is structured as follows. Section \ref{sec:3_related_works} analyzes the most common security problems in the mobile cloud environment and the solutions for offline protection in the BYOD world. Section \ref{sec:3_the_client_security_architecture} outlines the mobile security architecture of the proposed solution. Section \ref{sec:3_offline_mode} presents the detailed scheme of the proposed solution to the problem of offline mobile client security. Section \ref{sec:3_mos} explains the algorithms for MOS. Section \ref{sec:3_results} discusses the common threat scenarios, the data modeling, the complexity analysis and dicussion of the practical implementation results. Section \ref{sec:3_conclusion} concludes the chapter.

@@

In many cases the industrial providers of the mobile security API avoid the problem by completely forbidding the offline access to the protected app, i.e. SAP Mocona, which operates as a secure client wrapping layer \cite{Lelyveld2013}. Due to various constraints such as traffic load, travelling and ease of access, the SME business procedure may require such access.

To the best of our knowledge, the offline mode security problem has not yet been deployed, neither in academia nor in the industry \cite{yovel2014, khan2015cloud, khan2014survey, khan2013towards}. Therefore, the main concern of this proposal is the protection of the mobile client and its data in offline mode, when the functions of data protection cannot be offloaded to a cloud or a trusted party.

@@

The core set of functions and protocols of the proposed architecture can be divided into three sets of operations as shown in Figure \ref{fig:3_01}.

@@

Figure \ref{fig:3_02} depicts the client-side protection mechanisms. The client should support 4 subsystems: i) encryption subsystem (\textbf{Online/Offline Criptography}) that contains the procedures of encryption and decryption; ii) storage subsystem (\textbf{Protected Storage}) that provides the downloaded shares and key storage protection; iii) (\textbf{threat intelligence}) unit that provides the constant monitoring, and iv) the communication subsystem (\textbf{Communication with Server}). In short, all security procedures are connected to 4 groups of operations: file request and receiving; encryption and decryption; file and key storing; monitoring and analysis.

@@

This architecture consists of the modules of cryptographic functions, threat intelligence infrastructure, communication with server and storage. However, this work focus on the threat intelligence infrastructure module for malicious behavior analysis.

The threat intelligence infrastructure takes into account simple actors such as the time counter for the key expiry period, the counter of unsuccessful tries in order to protect from brute force attacks, and MOS-inspired statistics analyzer. Functions such as alerting and deleting the expired key belong to this module as well. These functions are described in Subsection \ref{sec:3_mos}.

@@

In the online mode the mobile client uses the secure communication with the server in order to verify the validity of user’s credentials. On the contrary, the offline protection model should be approached independently.

@@

The MOS schemes can rely on sample covariance of zero mean variables (called as \textbf{zero mean covariance} for the sake of simplicity) and sample covariance of zero mean and unitary standard deviation (called as \textbf{zero mean and standardized covariance} for the sake of simplicity) variables, where the former is useful to identify abnormalities caused by large amounts of operations during a period, while the latter is applied to identify anomalies on sparse or subtle number of file operations.

@@

Once the $\mathbf{\hat{R}}_{yy}$ or $\mathbf{\hat{R}}_{zz}$ have been obtained for MOS in order to anomaly detection, for the sake of simplicity, we refer to $\mathbf{\hat{R}}_{yy}$ or $\mathbf{\hat{R}}_{zz}$ as a matrix $\mathbf{C}$. Therefore, the next step of the algorithm is the eigenvalue decomposition (EVD), calculated according to $\mathbf{C}^{(q)} = \mathbf{V}^{(q)}\mathbf{\lambda}^{(q)}\mathbf{V}^{(q)^{\rm T}}$, in order to obtain the vector of eigenvalues $\textbf{e}$, as following:

\begin{equation}\label{eq:3_eq07}
	\mathbf{e}^{(q)} = \rm diag(\mathbf{\lambda}^{(q)}),
\end{equation}

@@

Note that some MOS schemes may also require the amount of time that compose a time frame, such as $\hat{d} = \textrm{MOS}(\boldsymbol{e}_{\rm max}, M)$. For more information about MOS schemes, interested readers are referred to Appendix \ref{apx:a_mos} and to \cite{da2009comparison}.

@@

This section provides the detailed analysis of the results from the proposed approach regarding security and complexity analysis.

@@

The session time defines the period when operations can be performed until the next session renewing. During this period, it is still necessary to identify attacks and malicious behavior on file operations, in order to avoid fast attacks to perform unauthorized access to information or data modification. Some attacks present behavioral patterns based on abrupt number of operations, such as the ransomware attack, which is a growing attack \cite{McAfee2015} that blocks the access to valuable resources and requires a payment in order to unblock the content. The access to the resources can be blocked by the attacker through some techniques, when the content is encrypted by the attacker, the ransomware attack can be called cryptoransomware or cryptolocker \cite{kaspersky2014}.

@@

\begin{enumerate}
	\item number of file decryption;
	\item number of decrypted file reading;
	\item number of decrypted file execution. 
\end{enumerate}

@@

\begin{enumerate}
	\item number of file encryption;
	\item number of decrypted file writing.
\end{enumerate}

@@

Table \ref{tab:3_04} presents the data modeling time and the processing time of eigenvalues decomposition calculations to be applied to anomaly detection from user operation logs of Storgrid mobile client.

@@

The information presented by column are the device model, the log size in megabytes, the window size in minutes, the data modeling time in milliseconds, the average of eigenvalue decomposition time in milliseconds, the standard deviation of eigenvalue decomposition time in milliseconds, the minimum of eigenvalue decomposition time in milliseconds and the maximum of eigenvalue decomposition time in milliseconds.

@@

Future works in the area can further explore enhancements in the analytics methods as well as to extended the approach to be evaluate multidimensional data through tensor-based analysis.
